Cell Decompositions for Rank Two Quiver Grassmannians

### 2019-06-10 Grassmannian_cells.tex
Section 3 edits

To avoid such ambiguities, we introduce the set
\[A_1^{(k)}:=\{(i_1,\ldots,i_k)\mid i_j\ne i_{j+1}\text{ for $1\le j\le k-1$}\}\subset\{1,\ldots,n\}^k\]
and call the elements of $A_1^{(k)}$ \emph{admissible sequences} of length $k$.
The following gives an analogue of Lemma~\ref{cor:truncated preprojective isomorphism}.
For $I\subsetneq\{1,\ldots,n\}$, write $I^c=\{1,\ldots,n\}\setminus I$ for the complementary subset.
\begin{lemma}
  \label{le:truncated quotients}
  For $m\ge3$, the following hold:
  \begin{enumerate}
    \item Given a decomposition $I\sqcup\{i\}=\{1,\ldots,n\}$ with $|I|=n-1$, there is a canonical isomorphism 
      \[\tilde P_{m+1}^I\cong\tilde P_{m,i}^{\{i\}}=\tilde P_{m,i}/\tilde P_{m-1,(i,i)}.\]
    \item For any pair $i,j\in\{1,\ldots,n\}$ with $i\ne j$ and any $J\subsetneq\{i\}^c$, there is a commutative diagram
      \begin{equation}
        \label{dia:truncated coherence}
        \xymatrix@R40pt@C60pt{& \tilde P_{m+1} \ar_(.45){\tilde\pi_{m+1}^J}[dl] \ar^(.4){\tilde\pi_{m+1}^{\{i\}^c}}[d] & \ar[l] \tilde P_{m,i} \ar_(.45){\tilde\pi_{m,i}^{\{i\}}}[dl]\ar^(.4){\tilde\pi_{m,i}^{\{j\}^c}}[d] & \ar[l] \tilde P_{m-1,(i,j)} \ar_(.45){\tilde\pi_{m-1,(i,j)}^{\{j\}}}[dl]\\
        \tilde P_{m+1}^J \ar^{\tilde\pi_{m+1}^{J,\{i\}^c}}[r] & \tilde P_{m+1}^{\{i\}^c}\cong\tilde P_{m,i}^{\{i\}}\ar^{\tilde\pi_{m,i}^{\{i\},\{j\}^c}}[r] & \tilde P_{m,i}^{\{j\}^c}\cong \tilde P_{m-1,(i,j)}^{\{j\}}}
      \end{equation}
    \item For any admissible word $\ui\in A_1^{(k)}$ with $1\le k\le m$ and any proper subsets $I,J\subsetneq\{1,\ldots,n\}$ with $i_k\in I$ and $i_1\notin J$, there exist (unique up to scale) surjective maps $\tilde\pi_{m+1}^{\ui,J,I}:\tilde P_{m+1}^J\to \tilde P_{m+1-k,\ui}^I$ satisfying:
      \[\tilde\pi_{m+1}^{\ui,J,I}\Big|_{\tilde P_{m+1-\ell,\ui_\ell}}=\tilde\pi_{m+1-\ell,\ui_\ell}^{\ui'_\ell,I}
      \qquad\text{and}\qquad
    \tilde\pi_{m+1}^{\ui,J,I}\Big|_{\tilde P_{m+1-k,\ui}}=\tilde\pi_{m+1-k,\ui}^I\]
      for $\ell<k$, where $\ui_\ell=(i_1,\ldots,i_\ell)$ and $\ui'_\ell=(i_{\ell+1},\ldots,i_k)$.
  \end{enumerate}
\end{lemma}
\begin{remark}
  Note that we may take $J=\varnothing$ in Lemma~\ref{le:truncated quotients}.3 but the conditions impose $I\ne\varnothing$.
\end{remark}
\begin{proof}
  The Auslander-Reiten sequence~\eqref{eq:lifted AR sequence} gives rise to the following commutative diagram:
  \[\xymatrix{& & 0\ar[d]& 0\ar[d]&\\
    & & \bigoplus_{i\in I} \tilde P_{m,i}\ar[d]\ar@{=}[r]& \bigoplus_{i\in I} \tilde P_{m,i}\ar[d]&\\
    0\ar[r]&\tilde P_{m-1} \ar[r]\ar@{=}[d] & \bigoplus_{i=1}^n \tilde P_{m,i}\ar[d]\ar[r]\ar@{}[dr]|(.7){\lrcorner} & \tilde P_{m+1}\ar[r]\ar[d]& 0\\
    0\ar[r] &\tilde P_{m-1}\ar[r] &  \tilde P_{m,j}\ar[r]\ar[d] & \tilde P_{m+1}^I\ar[r]\ar[d] & 0\\
  & & 0& 0&}\]
  The image of the inclusion $\tilde P_{m-1}\into \tilde P_{m,j}$ in the bottom row is the subrepresentation $\tilde P_{m-1,(j,j)}\subset\tilde P_{m,j}$ and part (1) follows.

@@

  These diagrams stitch together to give the following composition, where we suppress the lower isomorphisms in the diagram \eqref{dia:truncated coherence}:
  \[\tilde\pi_{m+1}^{\ui,J,I}:=\tilde\pi_{m+1-k,\ui}^{\{i_k\},I}\circ\tilde\pi_{m+2-k,\ui_{k-1}}^{\{i_{k-1}\},\{i_k\}^c}\circ\cdots\circ\tilde\pi_{m,i_0}^{\{i_1\},\{i_2\}^c}\circ\tilde\pi_{m+1}^{J,\{i_1\}^c}\]
  which can be seen to satisfy the desired properties from (3) inductively using the commutativity from (2).
\qed\end{proof}

@@

\begin{theorem}
  \label{cellscover}
  For $m\geq 1$ and $J\subsetneq\{1,\ldots,n\}$, every quiver Grassmannian $\Gr^{\tilde Q}_\tbfe(\tilde P_{m+1}^J)$ admits a cell decomposition.
\end{theorem}
\begin{corollary}
  Every quiver Grassmannian of a lifted preprojective representation admits a cell decomposition.
\end{corollary}
\begin{proof}
  This is the special case $J=\varnothing$ in Theorem~\ref{cellscover}.
\end{proof}
\begin[Proof of Theorem~\ref{cellscover}]{proof}

  We work by induction on $m$.
  When $m=1$, the claim is trivial since in this case all quiver Grassmannians are points.
  We establish the result for $m\ge2$ by proving the following more general statement:\bigskip

@@

  For $E\in\Gr^{\tilde Q}_\tbfe(\tilde P_{m+1}^J)$ which is contained in such a cell, we have $\pi_{m+1}^{J,I}(E)\in C_k$.
  But then $\pi_{m+1}^{J,\bfI}(E)\ne0$ if and only if $\pi_{m+1}^{I,\bfI}(G)\ne0$ for all $G\in C_k$ and thus $\pi_{m+1}^{J,\bfI}(E')\ne0$ for all $E'$ in this cell $B_\ell\times C_k\times\AA^d$ of $\Gr^{\tilde Q}_\tbfe(\tilde P_{m+1}^J)$. This shows ($\dagger$) for $\pi_{m+1}^{J,\bfI}$.
\qed\end{proof}

>>>>

To avoid such ambiguities, we introduce the set
\[A_1^{(k)}:=\{(i_1,\ldots,i_k)\mid i_j\ne i_{j+1}\text{ for $1\le j\le k-1$}\}\subset\{1,\ldots,n\}^k\]
and call the elements of $A_1^{(k)}$ \emph{admissible sequences} of length $k$.

The following gives an analogue of Lemma~\ref{cor:truncated preprojective isomorphism}.
For $I\subsetneq\{1,\ldots,n\}$, write $I^c=\{1,\ldots,n\}\setminus I$ for the complementary subset.
\begin{lemma}
  \label{le:truncated quotients}
  For $m\ge2$, the following hold:
  \begin{enumerate}
    \item Given a decomposition $I\sqcup\{i\}=\{1,\ldots,n\}$ with $|I|=n-1$, there is a canonical isomorphism 
      \[\tilde P_{m+1}^I\cong\tilde P_{m,i}^{\{i\}}=\tilde P_{m,i}/\tilde P_{m-1,(i,i)}.\]
    \item For any pair $i,j\in\{1,\ldots,n\}$ with $i\ne j$ and any $J\subsetneq\{i\}^c$, there is a commutative diagram
      \begin{equation}
        \label{dia:truncated coherence}
        \xymatrix@R40pt@C60pt{& \tilde P_{m+1} \ar_(.45){\tilde\pi_{m+1}^J}[dl] \ar^(.4){\tilde\pi_{m+1}^{\{i\}^c}}[d] & \ar[l] \tilde P_{m,i} \ar_(.45){\tilde\pi_{m,i}^{\{i\}}}[dl]\ar^(.4){\tilde\pi_{m,i}^{\{j\}^c}}[d] & \ar[l] \tilde P_{m-1,(i,j)} \ar_(.45){\tilde\pi_{m-1,(i,j)}^{\{j\}}}[dl]\\
        \tilde P_{m+1}^J \ar^{\tilde\pi_{m+1}^{J,\{i\}^c}}[r] & \tilde P_{m+1}^{\{i\}^c}\cong\tilde P_{m,i}^{\{i\}}\ar^{\tilde\pi_{m,i}^{\{i\},\{j\}^c}}[r] & \tilde P_{m,i}^{\{j\}^c}\cong \tilde P_{m-1,(i,j)}^{\{j\}}}
      \end{equation}
    \item For an admissible sequence $\ui\in A_1^{(k)}$, $k<m$, and $J\subsetneq\{1,\ldots,n\}$, the restriction of $\tilde\pi_{m+1}^J:\tilde P_{m+1}\to\tilde P_{m+1}^J$ to $\tilde P_{m+1-k,\ui}$ is injective if and only if $i_1\notin J$.
    \item For any admissible word $\ui\in A_1^{(k)}$ with $1\le k< m$ and any proper subsets $I,J\subsetneq\{1,\ldots,n\}$ with $i_k\in I$ and $i_1\notin J$, there exist (unique up to scale) surjective maps $\tilde\pi_{m+1}^{\ui,J,I}:\tilde P_{m+1}^J\to \tilde P_{m+1-k,\ui}^I$ satisfying:
      \begin{itemize}
        \item $\tilde\pi_{m+1}^{\ui,J,I}\circ\tilde\pi_{m+1}^{K,J}=\tilde\pi_{m+1}^{\ui,K,I}$ for $K\subsetneq J$;
        \item $\tilde\pi_{m+1-k,\ui}^{I,H}\circ\tilde\pi_{m+1}^{\ui,J,I}=\tilde\pi_{m+1}^{\ui,J,H}$ for $I\subsetneq H\subsetneq\{1,\ldots,n\}$;
        \item $\tilde\pi_{m+1}^{\ui,J,I}\Big|_{\tilde P_{m+1-k,\ui}}=\tilde\pi_{m+1-k,\ui}^I$;
        \item $\tilde\pi_{m+1}^{\ui,J,I}\Big|_{\tilde P_{m+1-\ell,\ui_\ell}}=\tilde\pi_{m+1-\ell,\ui_\ell}^{\ui'_\ell,\varnothing,I}$ for $\ell<k$, where $\ui_\ell:=(i_1,\ldots,i_\ell)$ and $\ui'_\ell:=(i_{\ell+1},\ldots,i_k)$.
      \end{itemize}
  \end{enumerate}
\end{lemma}
\begin{proof}
  The Auslander-Reiten sequence~\eqref{eq:lifted AR sequence} gives rise to the following commutative diagram:
  \[\xymatrix{& & 0\ar[d]& 0\ar[d]&\\
    & & \bigoplus_{i\in I} \tilde P_{m,i}\ar[d]\ar@{=}[r]& \bigoplus_{i\in I} \tilde P_{m,i}\ar[d]&\\
    0\ar[r]&\tilde P_{m-1} \ar[r]\ar@{=}[d] & \bigoplus_{i=1}^n \tilde P_{m,i}\ar[d]\ar[r]\ar@{}[dr]|(.7){\lrcorner} & \tilde P_{m+1}\ar[r]\ar[d]& 0\\
    0\ar[r] &\tilde P_{m-1}\ar[r] &  \tilde P_{m,j}\ar[r]\ar[d] & \tilde P_{m+1}^I\ar[r]\ar[d] & 0\\
  & & 0& 0&}\]
  The image of the inclusion $\tilde P_{m-1}\into \tilde P_{m,j}$ in the bottom row is the subrepresentation $\tilde P_{m-1,(j,j)}\subset\tilde P_{m,j}$ and part (1) follows.

@@

  For (3), if $i_1\in J$, then $\tilde P_{m,i_1}\subset\ker\tilde\pi_{m+1}^J$ and so the same is true of $\tilde P_{m+1-k,\ui}$.
  By Lemma~\ref{le:properties} and the admissibility of $\ui$, the representation $\tilde P_{m-1,(i_1,i_2)}$ embeds into $\tilde P_{m,i_1}^{\{i_1\}}$ and hence the same is true of $\tilde P_{m+1-k,\ui}$.
  For $i_1\notin J$, part (2) shows that such embeddings factor through $\tilde P_{m+1}^J$ which gives (3).

  The diagrams from (2) stitch together to give the following composition, where we suppress the lower isomorphisms in the diagram \eqref{dia:truncated coherence}:
  \[\tilde\pi_{m+1}^{\ui,J,I}:=\tilde\pi_{m+1-k,\ui}^{\{i_k\},I}\circ\tilde\pi_{m+2-k,\ui_{k-1}}^{\{i_{k-1}\},\{i_k\}^c}\circ\cdots\circ\tilde\pi_{m,i_0}^{\{i_1\},\{i_2\}^c}\circ\tilde\pi_{m+1}^{J,\{i_1\}^c}.\]
  These can be seen to satisfy the desired properties from (4) inductively using Lemma~\ref{le:properties}.5 and the commutativity from (2).
\qed\end{proof}

@@

\begin{theorem}
  \label{cellscover}
  For each $m\geq 1$ and $J\subsetneq\{1,\ldots,n\}$, every quiver Grassmannian $\Gr^{\tilde Q}_\tbfe(\tilde P_{m+1}^J)$ admits a cell decomposition.
  These cell decompositions are compatible with the projection maps $\tilde\pi_{m+1}^{\ui,J,I}:\tilde P_{m+1}^J\to\tilde P_{m+1-k,\ui}^I$ from Lemma~\ref{le:truncated quotients} in the following sense:
  \begin{itemize}
    \item[($\dagger$)] for any $E\in\Gr^{\tilde Q}_\tbfe(\tilde P_{m+1}^J)$ such that $\tilde\pi_{m+1}^{\ui,J,I}(E)\neq 0$ for some admissible sequence $\bfi=(i_1,\ldots,i_k)$, $k<m$, with $i_1\notin J$ and subset $I\subsetneq\{1,\ldots,n\}$ with $i_k\in I$, we have $\tilde\pi_{m+1}^{\ui,J,I}(E')\neq 0$ for all $E'\in C_E$, where $C_E$ is the affine cell which contains $E$.
  \end{itemize}
\end{theorem}
\begin{corollary}
  Every quiver Grassmannian of a lifted preprojective representation admits a cell decomposition.
\end{corollary}
\begin{proof}
  This is the special case $J=\varnothing$ in Theorem~\ref{cellscover}.
\end{proof}
\subsubsection*{Proof of Theorem~\ref{cellscover}}
  We work by induction on $m$.
  When $m=1$, the claim is trivial since in this case all quiver Grassmannians are points.
  We establish the result for $m\ge2$ by proving the following more general statement:\bigskip

@@

  For $E\in\Gr^{\tilde Q}_\tbfe(\tilde P_{m+1}^J)$ which is contained in such a cell, we have $\pi_{m+1}^{J,I}(E)\in C_k$.
  But then $\pi_{m+1}^{J,\bfI}(E)\ne0$ if and only if $\pi_{m+1}^{I,\bfI}(G)\ne0$ for all $G\in C_k$ and thus $\pi_{m+1}^{J,\bfI}(E')\ne0$ for all $E'$ in this cell $B_\ell\times C_k\times\AA^d$ of $\Gr^{\tilde Q}_\tbfe(\tilde P_{m+1}^J)$. This shows ($\dagger$) for $\pi_{m+1}^{J,\bfI}$.
\qed
